Quantitative processing of broadband data as implemented in a scientific splitbeam echosounder

### 2022-05-09 Paper/Manuscript.tex
Updated figures

In the example a linear sweep enveloped by a Hanning window is implemented (Fig.~\ref{fi:ytx}), where the parameters are the initial frequency, the final frequency, the pulse duration, the sampling rate, and the proportion of the signal that is tapered in each end, respectively. A slope of 0 and .5 indicates no tapering and tapering across the whole signal, respectively.
\begin{figure}
\includegraphics[width=16cm]{Fig_ytx}
\caption{\label{fi:ytx} Linear chirp pulses from 34 kHz to 45 kHz with a slope of 0.057 (blue) and 0.5 (orange), as implemented in the EK80.}
\end{figure}

@@

Forward/aft and port/starboard phase angles of target echoes are estimated by combining the transducer half signals thus: 
%
\begin{eqnarray}
\label{eq:phase1}
y_\along(\samplesymt) & = & y_{\textrm{pc,fore}}(\samplesymt) y_{\textrm{pc,aft}}^*(\samplesymt), \\
y_\athw(\samplesymt) & = & y_{\textrm{pc,star}}(\samplesymt) y_{\textrm{pc,port}}^*(\samplesymt),
\end{eqnarray}
%
where $y_\along(\samplesymt)$ is the electrical angle along the minor axis of the transducer (positive in the forward direction when ship-mounted) and $y_\athw(\samplesymt)$ the electrical angle along the major axis of the transducer (positive to starboard when ship-mounted), where complex signals are represented in the form $e^{j 2\pi \freqsym \timesym}$, where $j = \sqrt{-1}$. The physical echo arrival angles ($\along$ and $\athw$) are then given by:
%
\begin{eqnarray}
\label{eq:phase2}
\along(\samplesymt) & = & \arcsin\left( \frac{\atan\left( \Im(y_\along(\samplesymt)), \Re(y_\along(\samplesymt) \right)}{\anglefalong}\right) \\
\athw(\samplesymt) & = & \arcsin\left( \frac{\atan\left( \Im(y_\athw(\samplesymt)), \Re(y_\athw(\samplesymt) \right)}{\anglefathw}\right),
\end{eqnarray}
%
where $\anglefalong$ and $\anglefathw$ are constants that convert from phase angles to physical echo arrival angles (Fig.~\ref{fi:theta_phi}) and are derived from the transducer geometry and, $\fc$, the centre frequency of the chirp pulse \citep{ehrenberg1979}. The inverse sine is indicated by $\arcsin$,the four quadrant inverse tangent which returns values in the interval $[-\pi, \pi]$ inclusive is indicated by $\atan$, the real part of a complex number by $\Re$ and the imaginary part by $\Im$. As a mnemonic, the horizontal line in the symbol used for the forward/aft direction, $\along$, represents the pivot axis for the alongship angles and the near-vertical line in the $\athw$ symbol indicates the pivot axis for port/starboard angles.
%
\begin{figure}
\includegraphics[width=16cm]{Fig_theta_phi}
\caption{\label{fi:theta_phi}Physical angles $\along$ and $\athw$.}
\end{figure}
\clearpage
\section{Target strength}

@@

\begin{figure}
\includegraphics[width=16cm]{Fig_TS_echogram}
\caption{\label{Fig_TS_echogram} Sv as a function of ping number and range ($n$) for a calibration sphere. The sphere is located at approximately 6 m range. The red vertical line indicates the ping that is used for illustraing the $\ts(\freqsym)$ processing.}
\end{figure}

@@

\begin{figure}
\includegraphics[width=16cm]{Fig_Sv_echogram}
\caption{\label{Fig_Sv_echogram} Sv as a function of ping number and range range ($\samplesymt$) for a echogram raw data file. The upper yellow area shows the transmit pulse, a fish school is seen as a registration between 15 and 35 m range, and the sea flor is seen at approximately 50 m. The red vertical line indicates the ping that is used for illustraing the $\sv(\freqsym)$ processing.}
\end{figure}

>>>>

In the example a linear sweep enveloped by a Hanning window is implemented (Fig.~\ref{fi:ytx}), where the parameters are the initial frequency, the final frequency, the pulse duration, the sampling rate, and the proportion of the signal that is tapered in each end, respectively. A slope of 0 and .5 indicates no tapering and tapering across the whole signal, respectively.
\begin{figure}
\includegraphics[width=16cm]{Fig_ytx}
\caption{\label{fi:ytx} Envelope of linear chirp pulses from 92 kHz to 158 kHz with a slope of 0.057 (blue) and 0.5 (orange).}
\end{figure}

@@

Forward/aft and port/starboard phase angles of target echoes are estimated by combining the transducer half signals thus: 
%
\begin{eqnarray}
\label{eq:phase1}
y_\along(\samplesymt) & = & y_{\textrm{pc,fore}}(\samplesymt) y_{\textrm{pc,aft}}^*(\samplesymt), \\
y_\athw(\samplesymt) & = & y_{\textrm{pc,star}}(\samplesymt) y_{\textrm{pc,port}}^*(\samplesymt),
\end{eqnarray}
%
where $y_\along(\samplesymt)$ is the electrical angle along the minor axis of the transducer (positive in the forward direction when ship-mounted) and $y_\athw(\samplesymt)$ the electrical angle along the major axis of the transducer (positive to starboard when ship-mounted), where complex signals are represented in the form $e^{j 2\pi \freqsym \timesym}$, where $j = \sqrt{-1}$. The physical echo arrival angles ($\along$ and $\athw$) are then given by:
%
\begin{eqnarray}
\label{eq:phase2}
\along(\samplesymt) & = & \arcsin\left( \frac{\atan\left( \Im(y_\along(\samplesymt)), \Re(y_\along(\samplesymt) \right)}{\anglefalong}\right) \\
\athw(\samplesymt) & = & \arcsin\left( \frac{\atan\left( \Im(y_\athw(\samplesymt)), \Re(y_\athw(\samplesymt) \right)}{\anglefathw}\right),
\end{eqnarray}
%
where $\anglefalong$ and $\anglefathw$ are constants that convert from phase angles to physical echo arrival angles (Fig.~\ref{fi:theta_phi}) and are derived from the transducer geometry and, $\fc$, the centre frequency of the chirp pulse \citep{ehrenberg1979}. The inverse sine is indicated by $\arcsin$,the four quadrant inverse tangent which returns values in the interval $[-\pi, \pi]$ inclusive is indicated by $\atan$, the real part of a complex number by $\Re$ and the imaginary part by $\Im$. As a mnemonic, the horizontal line in the symbol used for the forward/aft direction, $\along$, represents the pivot axis for the alongship angles and the near-vertical line in the $\athw$ symbol indicates the pivot axis for port/starboard angles.
%
\begin{figure}
\includegraphics[width=16cm]{Fig_theta_phi}
\caption{\label{fi:theta_phi}The physical angles $\along$ and $\athw$ for the target strength example data (see below). The single target can be seen around sample number 1000 where the angles are less variable.}
\end{figure}
\clearpage
\section{Target strength}

@@

\begin{figure}
\includegraphics[width=16cm]{Fig_TS_echogram}
\caption{\label{Fig_TS_echogram} Sv as a function of ping number and range for a calibration sphere. The sphere is located at approximately 6 m range. The red vertical line indicates the ping that is used for illustrating the $\ts(\freqsym)$ processing.}
\end{figure}

@@

\begin{figure}
\includegraphics[width=16cm]{Fig_Sv_echogram}
\caption{\label{Fig_Sv_echogram} Sv as a function of ping number and range range ($\samplesymt$) for the raw data file used in the Sv(f) example. The upper yellow area shows the transmit pulse, a fish school is seen as a registration between 15 and 35 m range, and the sea floor is seen at approximately 50 m. The red vertical line indicates the ping that is used for illustrating the $\sv(\freqsym)$ processing.}
\end{figure}
